### 2024-01-21 outputs/paper/paper.qmd
Create basic structure for paper

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Yuanyi (Leo) Liu
thanks: "Code and data are available at: https://github.com/leoyliu/STA302-Paper-1."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

install.packages("knitr")
install.packages("janitor")
install.packages("lubridate")
install.packages("opendatatoronto")
install.packages("tidyverse")
install.packages("palmerpenguins")

library(knitr)
library(janitor)
library(lubridate)
library(opendatatoronto)
library(tidyverse)
library(palmerpenguins)
```

# Introduction

You can and should cross-reference sections and sub-sections.

The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

Talk way more about it.

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

\newpage

# References

### 2024-01-21 scripts/00-simulate_data.R

@@ -1,0 +1,41 @@
+#### Preamble ####
+# Purpose: Get data on 2021 shelter usage and make table
+# Author: Rohan Alexander
+# Email: rohan.alexander@utoronto.ca
+# Date: 1 July 2022
+# Prerequisites: -
+
+#### Workspace setup ####
+install.packages("knitr")
+install.packages("janitor")
+install.packages("lubridate")
+install.packages("opendatatoronto")
+install.packages("tidyverse")
+
+library(knitr)
+library(janitor)
+library(lubridate)
+library(opendatatoronto)
+library(tidyverse)
+
+#### Simulate ####
+
+set.seed(853)
+
+simulated_occupancy_data <-
+  tibble(
+    date = rep(x = as.Date("2021-01-01") + c(0:364), times = 3),
+    # Based on Eddelbuettel: https://stackoverflow.com/a/21502386
+    shelter = c(
+      rep(x = "Shelter 1", times = 365),
+      rep(x = "Shelter 2", times = 365),
+      rep(x = "Shelter 3", times = 365)
+    ),
+    number_occupied =
+      rpois(
+        n = 365 * 3,
+        lambda = 30
+      ) # Draw 1,095 times from the Poisson distribution
+  )
+
+head(simulated_occupancy_data)

### 2024-01-21 scripts/01-download_data.R

@@ -1,0 +1,18 @@
+#### Acquire ####
+toronto_shelters <-
+  # Each package is associated with a unique id  found in the "For 
+  # Developers" tab of the relevant page from Open Data Toronto
+  # https://open.toronto.ca/dataset/daily-shelter-overnight-service-occupancy-capacity/
+  list_package_resources("21c83b32-d5a8-4106-a54f-010dbe49f6f2") |>
+  # Within that package, we are interested in the 2021 dataset
+  filter(name == 
+           "daily-shelter-overnight-service-occupancy-capacity-2021.csv") |>
+  # Having reduced the dataset to one row we can get the resource
+  get_resource()
+
+write_csv(
+  x = toronto_shelters,
+  file = "toronto_shelters.csv"
+)
+
+head(toronto_shelters)

### 2024-01-23 outputs/paper/paper.qmd
Update paper content (first edition)

@@
 ---
-title: "My title"
-subtitle: "My subtitle if needed"
+title: "Shifting Shadows of the Six: Decoding Toronto's Crime Trends from 2014 to 2023"
 author: 
@@
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "In this study, we conducted a comprehensive analysis of crime trends in Toronto, utilizing a dataset detailing various criminal activities from 2014 to 2023.
+Our investigation focused on the evolution of different types of crimes, with a specific emphasis on contrasting violent crimes, such as assault and robbery, against non-violent offenses like theft and fraud.
+The findings reveal significant patterns in both categories, highlighting an evolving landscape of criminal behavior in Toronto.
+This research is crucial as it sheds light on the changing dynamics of urban crime, providing valuable insights for policymakers, law enforcement agencies, and community leaders in developing targeted strategies for crime prevention and enhancing public safety in metropolitan areas."
 format: pdf
@@
 
-install.packages("knitr")
-install.packages("janitor")
-install.packages("lubridate")
-install.packages("opendatatoronto")
-install.packages("tidyverse")
-install.packages("palmerpenguins")
+# install.packages("knitr")
+# install.packages("janitor")
+# install.packages("lubridate")
+# install.packages("opendatatoronto")
+# install.packages("tidyverse")
+# install.packages("palmerpenguins")
+# install.packages("here")
 
@@
 library(palmerpenguins)
+library(here)
 ```
 
-# Introduction
+# Introduction {#sec-intro}
 
-You can and should cross-reference sections and sub-sections.
+Urban crime, a multifaceted issue impacting communities worldwide, poses significant challenges to public safety and social stability.
+Toronto, Canada's largest city, is no exception.
+The broader context of urban crime in Toronto is critical in understanding its impact on the city's development, public perception, and policy formulation.
+Despite extensive research in urban criminology, there remains a gap in comprehensively understanding how different types of crimes, especially violent and non-violent, have evolved over recent years in Toronto.
+This gap is crucial for developing effective crime prevention strategies and enhancing public safety.
 
-The remainder of this paper is structured as follows. @sec-data....
+This paper presents a detailed analysis of crime trends in Toronto from 2014 to 2023.
+The study focuses on two primary crime categories: violent crimes, such as assault and robbery, and non-violent crimes, including theft and fraud.
+By leveraging a comprehensive dataset, we delve into how these crime categories have changed over time and their implications for urban safety and policing strategies.
+The analysis not only uncovers patterns and trends in these crime categories but also offers a comparative perspective, highlighting the distinct trajectories of violent and non-violent crimes.
+
+The findings of this study are significant for several reasons.
+First, they provide a nuanced understanding of the crime dynamics in a major urban center, offering insights that are vital for policymakers, law enforcement, and community leaders.
+The patterns observed suggest shifts in criminal behavior and possibly reflect broader social and economic changes in the city.
+Second, the comparative analysis between violent and non-violent crimes offers a unique perspective on the nature of criminal activities, contributing to a more targeted approach in crime prevention and community safety initiatives.
+
+The paper is structured to facilitate a comprehensive understanding of the study and its implications.
+Following this introduction @sec-intro, the second section presents the data @sec-data, detailing the data sources, analytical techniques, and the rationale behind the chosen methods.
+The third section discusses the results @sec-result, elaborating on the observed trends and patterns in crime data.
+The fourth section provides an in-depth discussion @sec-discussion of these findings, exploring potential factors influencing these trends and drawing connections to broader urban issues.
+Finally, the paper concludes with a summary of the findings, their implications for urban safety and policy, and suggestions for future research in this area.
 
@@
 
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
+## Overview
+This section @sec-data aims to offer an insightful understanding of the dataset utilized in our analysis, focusing on its content, origin, and the methods applied for data manipulation and visualization.
+
+## Source of the Data
+The dataset comprises crime statistics in Toronto, spanning from 2014 to 2023.
+It was sourced from Open Data Toronto[@cite1], a reputable provider of public safety and criminal statistics.
+This dataset is publicly available and is often used for academic and policy research due to its reliability and comprehensive coverage.
+
+## Alternative Dataset Consideration
+Although alternative datasets were considered, this specific dataset was selected for its detailed year-wise breakdown and the diversity of crime types it encompasses.
+Other datasets either lacked the temporal detail or the variety of crime categories present in our chosen dataset.
+
+## Data Manipulation and Cleaning in R [@citeR]
+The data was processed and cleaned using R, a powerful statistical programming language. Key operations included:
+
+* **Data Cleaning:** Addressing missing values, standardizing formats, and filtering out irrelevant variables (e.g., removing rate-based columns).
+
+* **Data Transformation:** Converting the dataset from a wide format to a long format for easier manipulation using the `tidyr` package.
+
+* **Aggregation:** Summarizing the data by year and crime type using the `dplyr` package.
+
+## Variables Explanation
+The dataset includes various crime categories, such as assault, theft, and other types, with each variable representing the annual count of reported cases for each crime type.
+This longitudinal data provides a detailed view of the crime trends in Toronto over a decade.
+
+TODO: include a summary chart later.
 
 ```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
+#| label: fig-trends-in-crime
+#| fig-cap: Trends in Crime Cases in Toronto (2014-2023)
 #| echo: false
 
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
+# Read the data
+crime_data <- read.csv(here("inputs/data/cleaned_toronto_crime_rates.csv"))
+
+# Remove unnecessary columns for this plot
+crime_data <- crime_data %>%
+  select(-grep("RATE", names(crime_data)))
+
+# Gather the data to a long format
+crime_data_long <- crime_data %>%
+  gather(key = "Crime_Year", value = "Count", -AREA_NAME)
+
+# Extract year and crime type from the 'Crime_Year' column
+crime_data_long <- crime_data_long %>%
+  separate(Crime_Year, into = c("Crime_Type", "Year"), sep = "_") %>%
+  mutate(Year = as.numeric(Year))
+
+# Aggregate the data by year and crime type
+crime_data_summary <- crime_data_long %>%
+  group_by(Crime_Type, Year) %>%
+  summarise(Total = sum(Count, na.rm = TRUE), .groups = "drop")
+
+# Plotting all crime types in one graph
+ggplot(crime_data_summary, aes(x = Year, y = Total, color = Crime_Type)) +
+  geom_line() +
+  labs(title = "Evolution of Various Crimes in Toronto", x = "Year", y = "Total Count") +
   theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
+  theme(legend.position="bottom") +
+  guides(color = guide_legend(title = "Crime Type"))
+```
+@fig-trends-in-crime reveals distinct trends for various crime types in Toronto over the years.
+We notice that assault crimes have been on the rise in recent years, this could signal a growing concern in public safety and might prompt law enforcement to investigate the underlying causes, such as increased urbanization, changes in population density, or other socio-economic factors.
+Conversely, there is a noticeable decline in theft-related crimes, it could reflect the success of recent crime prevention initiatives, improved security measures, or increased public awareness.
+
+## Data Visualization
+For visual representation, `ggplot2`, a versatile package in R for data visualization, was employed.
+It was used to create comprehensive line plots depicting the trends of various crime categories over the years.
+These plots are instrumental in visualizing the temporal patterns and changes in crime statistics.
+
+## Summary Statistics and Relationship Analysis
+The dataset was summarized using basic statistical measures (mean, median, standard deviation) to provide an overview of each crime category.
+Further, the relationships between different crime types were explored to understand correlated trends, if any.
+
+# Results {#sec-result}
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+
 ```
 
-Talk more about it.
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-modelresults
+#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
+#| warning: false
+```
 
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
+Based on the analysis of the provided dataset and the visual representation of crime trends in Toronto over the years, several key results can be summarized:
 
-Talk way more about it.
+* **Variation in Crime Trends:** The dataset reveals distinct trends for different types of crimes in Toronto.
+Some crime categories exhibit an upward trend, indicating an increase in incidents over the years, while others show a downward trend, suggesting a decrease in occurrences.
+This variation highlights the complexity of crime dynamics in an urban setting.
 
-# Discussion
+* **Potential Factors Influencing Trends:** The observed trends in crime rates could be influenced by a multitude of factors.
+These include socio-economic changes, urban development, law enforcement strategies, policy reforms, and technological advancements in crime reporting.
+For example, an increase in certain types of crimes might be linked to economic downturns, demographic shifts, or changes in urban density, while a decrease could reflect effective policing strategies or community intervention programs.
 
-## First discussion point {#sec-first-point}
+# Discussion {#sec-discussion}
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
+## Implications for Policymakers and Law Enforcement: {#sec-first-point}
+
+The findings of this analysis are particularly relevant for policymakers and law enforcement agencies.
+Understanding these trends is crucial for developing targeted strategies to address and prevent crime effectively, thereby enhancing public safety and community well-being in Toronto.
 
@@
 
-## Third discussion point
-
 ## Weaknesses and next steps
 
-Weaknesses and next steps should also be included.
-
-\newpage
-
-\appendix
+While the visual analysis provides a high-level overview of crime trends in Toronto, it's essential to conduct more detailed investigations to understand the underlying causes of these trends.
+This includes examining subcategories of crimes, focusing on specific geographic areas within the city, and considering other relevant data sources.
 

### 2024-01-23 scripts/00-simulate_data.R

@@ -1,9 +1,8 @@
 #### Preamble ####
 # Purpose: Get data on 2021 shelter usage and make table
-# Author: Rohan Alexander
-# Email: rohan.alexander@utoronto.ca
-# Date: 1 July 2022
-# Prerequisites: -
+# Author: Yuanyi (Leo) Liu
+# Email: leoy.liu@mail.utoronto.ca
+# Date: 21 January 2024
 
 #### Workspace setup ####
 install.packages("knitr")

### 2024-01-23 scripts/01-download_data.R

@@ -1,18 +1,30 @@
-#### Acquire ####
-toronto_shelters <-
-  # Each package is associated with a unique id  found in the "For 
-  # Developers" tab of the relevant page from Open Data Toronto
-  # https://open.toronto.ca/dataset/daily-shelter-overnight-service-occupancy-capacity/
-  list_package_resources("21c83b32-d5a8-4106-a54f-010dbe49f6f2") |>
-  # Within that package, we are interested in the 2021 dataset
-  filter(name == 
-           "daily-shelter-overnight-service-occupancy-capacity-2021.csv") |>
-  # Having reduced the dataset to one row we can get the resource
+#### Preamble ####
+# Purpose: Downloads, saves, and reads data from the OpenDataToronto portal
+# Author: Yuanyi (Leo) Liu
+# Date: 21 January 2024
+# Contact: leoy.liu@mail.utoronto.ca
+# License: MIT
+
+#### Workspace setup ####
+library(opendatatoronto)
+library(tidyverse)
+library(dplyr)
+library(readr)
+
+#### Download data ####
+# Each package is associated with a unique id  found in the "For 
+# Developers" tab of the relevant page from Open Data Toronto
+# source: https://open.toronto.ca/dataset/neighbourhood-crime-rates/
+toronto_crime_rates <-
+  list_package_resources("neighbourhood-crime-rates") |>
+  filter(name == "neighbourhood-crime-rates - 4326.csv") |>
   get_resource()
 
+
+#### Save data ####
 write_csv(
-  x = toronto_shelters,
-  file = "toronto_shelters.csv"
+  x = toronto_crime_rates,
+  file = "toronto_crime_rates.csv"
 )
 
-head(toronto_shelters)
+head(toronto_crime_rates)

### 2024-01-23 scripts/02-data_cleaning.R

@@ -1,11 +1,28 @@
-toronto_shelters_clean <-
-  clean_names(toronto_shelters) |>
-  mutate(occupancy_date = ymd(occupancy_date)) |> 
-  select(occupancy_date, occupied_beds)
+#### Preamble ####
+# Purpose: Cleans the raw plane data
+# Author: Yuanyi (Leo) Liu
+# Date: 21 January 2024
+# Contact: leoy.liu@mail.utoronto.ca
+# License: MIT
 
-head(toronto_shelters_clean)
+#### Workspace setup ####
+library(tidyverse)
 
+#### Clean data ####
+raw_data <- read_csv("inputs/data/toronto_crime_rates.csv")
+
+# Removing unwanted columns
+cleaned_data <- raw_data %>% select(-c(X_id, HOOD_ID, POPULATION_2023, geometry))
+
+# Removing rows with NA values
+cleaned_data <- na.omit(cleaned_data)
+
+#### Save data ####
 write_csv(
-  x = toronto_shelters_clean,
-  file = "cleaned_toronto_shelters.csv"
+  x = cleaned_data,
+  file = "cleaned_toronto_crime_rates.csv"
 )
+
+#### Descriptive Summaries ####
+str(cleaned_data)
+summary(cleaned_data)

### 2024-01-23 outputs/paper/paper.qmd
Added references

@@
 
-* **Data Transformation:** Converting the dataset from a wide format to a long format for easier manipulation using the `tidyr` package.
+* **Data Transformation:** Converting the dataset from a wide format to a long format for easier manipulation using the `tidyr`[@citetidyr] package.
 
-* **Aggregation:** Summarizing the data by year and crime type using the `dplyr` package.
+* **Aggregation:** Summarizing the data by year and crime type using the `dplyr`[@citedplyr] package.
 

### 2024-01-24 outputs/paper/paper.qmd
Update paper content, especially for data section and discussion section

@@
 abstract: "In this study, we conducted a comprehensive analysis of crime trends in Toronto, utilizing a dataset detailing various criminal activities from 2014 to 2023.
-Our investigation focused on the evolution of different types of crimes, with a specific emphasis on contrasting violent crimes, such as assault and robbery, against non-violent offenses like theft and fraud.
+Our investigation focused on the evolution of different types of crimes, with a specific emphasis on contrasting violent crimes, such as assault and robbery, against non-violent offenses like theft.
 The findings reveal significant patterns in both categories, highlighting an evolving landscape of criminal behavior in Toronto.
@@
 # install.packages("knitr")
-# install.packages("janitor")
-# install.packages("lubridate")
 # install.packages("opendatatoronto")
 # install.packages("tidyverse")
-# install.packages("palmerpenguins")
 # install.packages("here")
+# install.packages("tidyr")
+# install.packages("dplyr")
+# install.packages("kableExtra")
 
 library(knitr)
-library(janitor)
-library(lubridate)
 library(opendatatoronto)
 library(tidyverse)
-library(palmerpenguins)
 library(here)
+library(tidyr)
+library(dplyr)
+library(kableExtra)
+library(janitor)
 ```
@@
 The third section discusses the results @sec-result, elaborating on the observed trends and patterns in crime data.
-The fourth section provides an in-depth discussion @sec-discussion of these findings, exploring potential factors influencing these trends and drawing connections to broader urban issues.
-Finally, the paper concludes with a summary of the findings, their implications for urban safety and policy, and suggestions for future research in this area.
+The fourth section provides an in-depth discussion @sec-discussion of these findings, exploring potential factors influencing these trends, drawing connections to broader urban issues, and suggestions for future research in this area.
 
@@
 The dataset comprises crime statistics in Toronto, spanning from 2014 to 2023.
-It was sourced from Open Data Toronto[@cite1], a reputable provider of public safety and criminal statistics.
+The dataset analyzed in this report was obtained in csv format from the City of Toronto Open Data Portal[@citedata] using the R package `opendatatoronto`[@citeopendatapkg].
+The dataset was last updated on January 14th, 2024.
 This dataset is publicly available and is often used for academic and policy research due to its reliability and comprehensive coverage.
@@
 ## Data Manipulation and Cleaning in R [@citeR]
-The data was processed and cleaned using R, a powerful statistical programming language. Key operations included:
+The data was processed and cleaned using R[@citeR], a powerful statistical programming language. Key operations included:
 
-* **Data Cleaning:** Addressing missing values, standardizing formats, and filtering out irrelevant variables (e.g., removing rate-based columns).
+* **Data Cleaning:** Addressing missing values, standardizing formats, and filtering out irrelevant variables using the `tidyverse`[@citetidyverse] package(e.g., removing rate-based columns and NAs).
 
@@
 
+* **Data Visualization:** For visual representation, `ggplot2` [@citeggplot], a versatile package in R for data visualization, was employed.
+It was used to create comprehensive line plots depicting the trends of various crime categories over the years.
+
+* **Paper Generation:** The packages `knitr` [@citeknitr] is used to generate the R markdown report.
+
 ## Variables Explanation
@@
 
-TODO: include a summary chart later.
-
 ```{r}
-#| label: fig-trends-in-crime
-#| fig-cap: Trends in Crime Cases in Toronto (2014-2023)
+#| label: tabledatasummary
 #| echo: false
@@
 
+# Creating the summary table
+summary_table <- crime_data_summary %>%
+  group_by(Crime_Type, Year) %>%
+  summarise(Total_Cases = sum(Total, na.rm = TRUE), .groups = "drop") %>%
+  spread(key = Year, value = Total_Cases)
+
+# Define a mapping for crime types
+readable_crime_types <- c('ASSAULT' = 'Assault',
+                          'AUTOTHEFT' = 'Auto Theft',
+                          'BIKETHEFT' = 'Bike Theft',
+                          'BREAKENTER' = 'Break and Enter',
+                          'HOMICIDE' = 'Homicide',
+                          'ROBBERY' = 'Robbery',
+                          'SHOOTING' = 'Shooting Incidents',
+                          'THEFTFROMMV' = 'Theft from Motor Vehicles',
+                          'THEFTOVER' = 'Major Theft')
+
+# Update the Crime_Type column
+summary_table$Crime_Type <- readable_crime_types[summary_table$Crime_Type]
+
+# Make a beautiful table
+summary_table %>%
+  slice(1:10) %>%
+  kable(
+    caption = "Crime Incidents Summary by Year",
+    col.names = c("Crime Type", "2014", "2015", "2016", "2017", "2018", "2019", "2020", "2021", "2022", "2023"),
+    digits = 1,
+    booktabs = TRUE,
+    linesep = "",
+    align = c("l", "l", "c", "c", "r", "r"),
+    format.args = list(big.mark = ",")) %>%
+    kable_styling(latex_options = "HOLD_position", font_size = 10, position = "center")
+```
+Table 1 organizes data into a clear format where each row indicates a specific type of crime.
+The columns are segmented by year, spanning from 2014 to 2023.
+Each cell within the table provides a count of the incidents reported for that particular crime in the corresponding year.
+
+Then, we want to categorize our data into one of two groups: 'Violent' or 'Non-Violent'.
+This categorization was based on the nature of the crime.
+For example, crimes like 'ASSAULT', 'HOMICIDE', etc., were categorized as 'Violent', while others like 'THEFT', were categorized as 'Non-Violent'.
+After categorization, the dataset was aggregated based on these two new categories.
+This involved summing up the incident counts for each category (Violent and Non-Violent) for every year.
+
+```{r}
+#| label: combineddatasummary
+#| echo: false
+summary_table$Category <- ifelse(summary_table$Crime_Type %in% c("Assault", "Break and Enter", "Homicide", "Robbery", "Shooting Incidents"), "Violent", "Non-Violent")
+
+# Group data by the new Category and sum the cases for each year
+two_types_crime <- summary_table %>%
+  group_by(Category) %>%
+  summarise(across(`2014`:`2023`, ~sum(.x, na.rm = TRUE)))
+
+# Make a beautiful table
+two_types_crime %>%
+  slice(1:10) %>%
+  kable(
+    caption = "Crime Incidents Summary by Year",
+    col.names = c("Crime Type", "2014", "2015", "2016", "2017", "2018", "2019", "2020", "2021", "2022", "2023"),
+    digits = 1,
+    booktabs = TRUE,
+    linesep = "",
+    align = c("l", "l", "c", "c", "r", "r"),
+    format.args = list(big.mark = ",")) %>%
+    kable_styling(latex_options = "HOLD_position", font_size = 10, position = "center")
+```
+
+Table 2 organized view of crime statistics over a range of years, specifically differentiated into two main categories: 'Violent' and 'Non-Violent' crimes.
+Each row in the table represents one of these two categories.
+The columns of the table are divided by year, ranging from 2014 to 2023.
+Under each year's column, the table displays the total number of reported incidents for each crime category.
+For example, under the column for 2014, you will find the total number of violent crimes reported in that year followed by the total number of non-violent crimes for the same year.
+
+## Summary Statistics and Relationship Analysis
+The dataset was summarized using basic statistical measures (mean, median, standard deviation) to provide an overview of each crime category.
+Further, the relationships between different crime types were explored to understand correlated trends, if any.
+
+\newpage
+
+# Results {#sec-result}
+
+```{r}
+#| label: fig-trends-in-crime
+#| fig-cap: Trends in Crime Cases in Toronto (2014-2023)
+#| echo: false
+
 # Plotting all crime types in one graph
@@
 
-## Data Visualization
-For visual representation, `ggplot2`, a versatile package in R for data visualization, was employed.
-It was used to create comprehensive line plots depicting the trends of various crime categories over the years.
-These plots are instrumental in visualizing the temporal patterns and changes in crime statistics.
+```{r}
+#| label: fig-trends-in-two-types-crime
+#| fig-cap: Comparative Trend Analysis of Violent and Non-Violent Crime Cases from 2014 to 2023
+#| echo: false
 
-## Summary Statistics and Relationship Analysis
-The dataset was summarized using basic statistical measures (mean, median, standard deviation) to provide an overview of each crime category.
-Further, the relationships between different crime types were explored to understand correlated trends, if any.
+# Ensure the data is ungrouped
+two_types_crime <- ungroup(two_types_crime)
 
-# Results {#sec-result}
+# Pivot the data to long format for plotting
+# Explicitly select the year columns to avoid issues with hidden grouping columns
+long_data <- two_types_crime %>%
+  select(Category, `2014`:`2023`) %>%
+  pivot_longer(cols = -Category, names_to = "Year", values_to = "Cases")
 
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
+# Create the plot
+crime_plot <- ggplot(long_data, aes(x = Year, y = Cases, group = Category, color = Category)) +
+  geom_line(linewidth = 1) + # Use linewidth instead of size
+  geom_point(size = 3) + # Point size
+  theme_minimal() + # Minimal theme
+  labs(title = "Evolution of Violent and Non-Violent Crimes Over the Years",
+       x = "Year",
+       y = "Number of Cases") +
+  scale_color_manual(values = c("Violent" = "red", "Non-Violent" = "blue")) +
+  theme(
+    plot.title = element_text(hjust = 0.5), # Center the plot title
+    legend.title = element_blank(), # Remove the legend title
+    axis.text.x = element_text(angle = 45, hjust = 1) # Angle the x-axis text for readability
+  )
 
+# Print the plot
+print(crime_plot)
 ```
 
-```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
-#| warning: false
-```
+@fig-trends-in-two-types-crime compares the number of cases reported for violent and non-violent crimes from 2014 to 2023.
+The x-axis represents the years, and the y-axis indicates the number of cases.
+
+The red line represents violent crimes, and the blue line represents non-violent crimes. The plot shows distinct trends for both categories over the specified time frame.
+
+* **Violent Crimes (Red Line):** There is an overall increasing trend in the number of violent crime cases over the years.
+There are some fluctuations, with a noticeable dip occurring in one of the years, but the general direction is upwards, especially towards the end of the period, indicating a rise in violent crimes or an increase in reporting.
+
+* **Non-Violent Crimes (Blue Line):** The trend in non-violent crime cases shows some variability but does not have as clear an increasing pattern as violent crimes.
+There are years where the rate has increased or decreased, but the last year shows a notable increase, suggesting a possible upward trend at the end of the period.
+
+Overall, the graph indicates that while both violent and non-violent crimes have seen rises and falls, the increase in violent crimes towards the later years is more pronounced.
+This could reflect changes in societal conditions, law enforcement practices, or reporting mechanisms.
+
+# Discussion {#sec-discussion}
 
@@
 
-* **Variation in Crime Trends:** The dataset reveals distinct trends for different types of crimes in Toronto.
+## Variation in Crime Trends: {#sec-first-point}
+
+The dataset reveals distinct trends for different types of crimes in Toronto.
 Some crime categories exhibit an upward trend, indicating an increase in incidents over the years, while others show a downward trend, suggesting a decrease in occurrences.
@@
 
-* **Potential Factors Influencing Trends:** The observed trends in crime rates could be influenced by a multitude of factors.
+## Potential Factors Influencing Trends: {#sec-second-point}
+
+The observed trends in crime rates could be influenced by a multitude of factors.
 These include socio-economic changes, urban development, law enforcement strategies, policy reforms, and technological advancements in crime reporting.
@@
 
-# Discussion {#sec-discussion}
-
-## Implications for Policymakers and Law Enforcement: {#sec-first-point}
-
 The findings of this analysis are particularly relevant for policymakers and law enforcement agencies.
 Understanding these trends is crucial for developing targeted strategies to address and prevent crime effectively, thereby enhancing public safety and community well-being in Toronto.
-
-## Second discussion point
 

### 2024-01-24 scripts/01-download_data.R

@@ -9,7 +9,6 @@
 library(opendatatoronto)
 library(tidyverse)
 library(dplyr)
-library(readr)
 
 #### Download data ####
 # Each package is associated with a unique id  found in the "For 

### 2024-01-25 outputs/paper/paper.qmd
Added test and some polish for the paper

--- 2024-01-23 outputs/paper/paper.qmd
@@
 abstract: "In this study, we conducted a comprehensive analysis of crime trends in Toronto, utilizing a dataset detailing various criminal activities from 2014 to 2023.
-Our investigation focused on the evolution of different types of crimes, with a specific emphasis on contrasting violent crimes, such as assault and robbery, against non-violent offenses like theft and fraud.
+Our analysis focused on the evolution of different types of crimes, with a specific emphasis on contrasting violent crimes, such as assault and robbery, against non-violent offenses like theft.
 The findings reveal significant patterns in both categories, highlighting an evolving landscape of criminal behavior in Toronto.
@@
 format: pdf
+toc: true
 number-sections: true
@@
 
-Urban crime, a multifaceted issue impacting communities worldwide, poses significant challenges to public safety and social stability.
+Urban crime, a significant issue impacting communities worldwide, poses substantial threats to public safety and social stability.
 Toronto, Canada's largest city, is no exception.
@@
 
-This paper presents a detailed analysis of crime trends in Toronto from 2014 to 2023.
-The study focuses on two primary crime categories: violent crimes, such as assault and robbery, and non-violent crimes, including theft and fraud.
-By leveraging a comprehensive dataset, we delve into how these crime categories have changed over time and their implications for urban safety and policing strategies.
+This paper provides a thorough examination of Toronto crime patterns from 2014 to 2023.
+The paper focuses on two types of crimes: violent crimes and nonviolent crimes.
+By using a comprehensive dataset, we delve into how these crime categories have changed over time and their implications for urban safety and policing strategies.
 The analysis not only uncovers patterns and trends in these crime categories but also offers a comparative perspective, highlighting the distinct trajectories of violent and non-violent crimes.
 
-The findings of this study are significant for several reasons.
-First, they provide a nuanced understanding of the crime dynamics in a major urban center, offering insights that are vital for policymakers, law enforcement, and community leaders.
+The findings of this paper are noteworthy for several reasons.
+First, they provide a nuanced understanding of the crime dynamics a large metropolitan area, offering insights that are vital for policymakers, law enforcement, and community leaders.
 The patterns observed suggest shifts in criminal behavior and possibly reflect broader social and economic changes in the city.
@@
 The paper is structured to facilitate a comprehensive understanding of the study and its implications.
-Following this introduction @sec-intro, the second section presents the data @sec-data, detailing the data sources, analytical techniques, and the rationale behind the chosen methods.
-The third section discusses the results @sec-result, elaborating on the observed trends and patterns in crime data.
-The fourth section provides an in-depth discussion @sec-discussion of these findings, exploring potential factors influencing these trends and drawing connections to broader urban issues.
-Finally, the paper concludes with a summary of the findings, their implications for urban safety and policy, and suggestions for future research in this area.
+Following @sec-intro, @sec-data presents the data, detailing the data sources, analytical techniques, and the rationale behind the chosen methods.
+@sec-result discusses the results, elaborating on the observed trends and patterns in crime data.
+@sec-discussion provides an in-depth discussion of these findings, exploring potential factors influencing these trends, drawing connections to broader urban issues, and suggestions for future research in this area.
+
+\newpage
 
@@
 
-## Overview
-This section @sec-data aims to offer an insightful understanding of the dataset utilized in our analysis, focusing on its content, origin, and the methods applied for data manipulation and visualization.
+This section aims to offer an insightful understanding of the dataset utilized in our analysis, focusing on its content, origin, and the methods applied for data manipulation and visualization.
 
@@
 The dataset comprises crime statistics in Toronto, spanning from 2014 to 2023.
-It was sourced from Open Data Toronto[@cite1], a reputable provider of public safety and criminal statistics.
-This dataset is publicly available and is often used for academic and policy research due to its reliability and comprehensive coverage.
+The dataset analyzed in this report was obtained in csv format from the City of Toronto Open Data Portal[@citedata] using the R package `opendatatoronto`[@citeopendatapkg].
+It includes the Crime Data by Neighbourhood.
+Counts are available for Assault, Auto Theft, Break and Enter, Robbery, Theft Over, Homicide and Shooting & Firearm Discharges.
+Data also includes the crime rate per 100,000 population calculated using the population estimates provided by Environics Analytics.
+For the purpose of this study, we focus solely on the number of reported incidents.
+The data, which is updated as of January 14th, 2024, is a well-regarded resource for scholarly and policy-making endeavors, valued for its thoroughness and dependability.
 
-## Alternative Dataset Consideration
 Although alternative datasets were considered, this specific dataset was selected for its detailed year-wise breakdown and the diversity of crime types it encompasses.
@@
 
-## Data Manipulation and Cleaning in R [@citeR]
-The data was processed and cleaned using R, a powerful statistical programming language. Key operations included:
+## Data Manipulation and Cleaning
+The data was processed and cleaned using R[@citeR], a powerful statistical programming language. Key operations included:
 
-* **Data Cleaning:** Addressing missing values, standardizing formats, and filtering out irrelevant variables (e.g., removing rate-based columns).
+* **Data Cleaning:** The initial phase of cleaning the Toronto crime rates dataset involved importing the raw data using the `read_csv` function from the `readr`[@citereadr] package.
+Following the import, the dataset was streamlined using the `select` function from `dplyr`[@citedplyr], a tidyverse[@citetidyverse] package, to remove irrelevant columns such as `X_id`, `HOOD_ID`, `POPULATION_2023`, and `geometry`.
+This step was crucial to focus the analysis on relevant variables.
+Additionally, the `na.omit` function was employed to discard any rows with missing values (NA), ensuring the dataset's completeness and reliability.
+The final step in the cleaning process was the exportation of the cleaned data into a new CSV file using the `write_csv `function from `readr`[@citereadr], thereby preserving the cleaned and refined dataset for subsequent analysis.
 
-* **Data Transformation:** Converting the dataset from a wide format to a long format for easier manipulation using the `tidyr` package.
+* **Data Transformation:** The dataset was transformed from a wide format to a long format using the `gather` function from the `tidyr`[@citetidyr] package.
+This transformation is crucial for simplifying the data structure and making it more amenable to analysis.
+Also, the separate function, again from `tidyr`[@citetidyr], was instrumental in dividing the 'Crime_Year' column into two distinct parts: 'Crime_Type' and 'Year'.
+This enhances the granularity of the dataset, allowing for more detailed analysis.
 
-* **Aggregation:** Summarizing the data by year and crime type using the `dplyr` package.
+* **Aggregation:** The `group_by` and `summarise` functions from `dplyr`[@citedplyr] were employed to aggregate the data by year and crime type.
+
+* **Data Visualization:** For visual representation, `ggplot2`[@citeggplot], a versatile package in R for data visualization, was employed.
+It was used to create comprehensive line plots depicting the trends of various crime categories over the years.
+
+* **Paper Generation:** The packages `knitr`[@citeknitr] is used in generating R Markdown reports.
 
@@
 
-TODO: include a summary chart later.
-
 ```{r}
-#| label: fig-trends-in-crime
-#| fig-cap: Trends in Crime Cases in Toronto (2014-2023)
+#| label: tbl-data-summary
+#| tbl-cap: Crime Incidents Summary by Year
 #| echo: false
@@
 # Read the data
-crime_data <- read.csv(here("inputs/data/cleaned_toronto_crime_rates.csv"))
+summary_table <- read.csv(here("inputs/data/summary_table.csv"), check.names = FALSE)
 
-# Remove unnecessary columns for this plot
-crime_data <- crime_data %>%
-  select(-grep("RATE", names(crime_data)))
+# Make a summary table
+summary_table %>%
+  slice(1:10) %>%
+  kable(
+    col.names = c("Crime Type", "2014", "2015", "2016", "2017", "2018", "2019", "2020", "2021", "2022", "2023"),
+    digits = 1,
+    booktabs = TRUE,
+    linesep = "",
+    align = c("l", "l", "c", "c", "r", "r"),
+    format.args = list(big.mark = ",")) %>%
+    kable_styling(latex_options = "HOLD_position", font_size = 10, position = "center")
+```
+For @tbl-data-summary, we observe a detailed enumeration of various crime incidents reported annually in Toronto from 2014 to 2023.
+Assault cases dominate the data, climbing substantially in subsequent years, indicating a worsening public safety concern.
+Each crime category follows its own trajectory, with theft from motor vehicles and major theft also showing significant fluctuations.
 
-# Gather the data to a long format
-crime_data_long <- crime_data %>%
-  gather(key = "Crime_Year", value = "Count", -AREA_NAME)
+We proceeded to organize the dataset into two categories: 'Violent' crimes, which included offenses such as 'Assault' and 'Homicide', and 'Non-Violent' crimes, encompassing acts like 'Theft'.
+Following this classification, we combined the data into these new groupings, totaling the number of incidents in each category for each year of the research.
+```{r}
+#| label: tbl-combined-data-summary
+#| tbl-cap: Crime Incidents Summary of Violent and Non-Violent Crime Cases by Year
+#| echo: false
 
-# Extract year and crime type from the 'Crime_Year' column
-crime_data_long <- crime_data_long %>%
-  separate(Crime_Year, into = c("Crime_Type", "Year"), sep = "_") %>%
-  mutate(Year = as.numeric(Year))
+# Read the data
+two_types_crime <- read.csv(here("inputs/data/two_types_crime.csv"), check.names = FALSE)
 
-# Aggregate the data by year and crime type
-crime_data_summary <- crime_data_long %>%
-  group_by(Crime_Type, Year) %>%
-  summarise(Total = sum(Count, na.rm = TRUE), .groups = "drop")
+# Make a table
+two_types_crime %>%
+  slice(1:10) %>%
+  kable(
+    col.names = c("Crime Type", "2014", "2015", "2016", "2017", "2018", "2019", "2020", "2021", "2022", "2023"),
+    digits = 1,
+    booktabs = TRUE,
+    linesep = "",
+    align = c("l", "l", "c", "c", "r", "r"),
+    format.args = list(big.mark = ",")) %>%
+    kable_styling(latex_options = "HOLD_position", font_size = 10, position = "center")
+```
 
-# Plotting all crime types in one graph
-ggplot(crime_data_summary, aes(x = Year, y = Total, color = Crime_Type)) +
-  geom_line() +
-  labs(title = "Evolution of Various Crimes in Toronto", x = "Year", y = "Total Count") +
-  theme_minimal() +
-  theme(legend.position="bottom") +
-  guides(color = guide_legend(title = "Crime Type"))
+@tbl-combined-data-summary organized view of crime statistics over a range of years, specifically differentiated into two main categories: 'Violent' and 'Non-Violent' crimes.
+Each row in the table represents one of these two categories.
+The columns of the table are divided by year, ranging from 2014 to 2023.
+Under each year's column, the table displays the total number of reported incidents for each crime category.
+Here, violent crimes exhibit a steady rise over the years, dwarfing the non-violent crimes, which display slight variability.
+
+\newpage
+
+# Results {#sec-result}
+In this section, we delve into the core findings of our study, shedding light on the intricate patterns and trends of crime in Toronto from 2014 to 2023.
+Our analysis dissects both violent and non-violent crime statistics, revealing insightful correlations and variations over the decade.
+```{r fig.width=10, fig.height=5.5}
+#| label: fig-trends-in-crime
+#| fig-cap: Trends in Crime Cases in Toronto
+#| echo: false
+
+# Transform the data into a long format
+long_data <- gather(summary_table, Year, Count, -Crime_Type)
+
+# Plotting the line graph
+ggplot(long_data, aes(x = Year, y = Count, group = Crime_Type, color = Crime_Type)) +
+  geom_line(linewidth = 1) + # Use linewidth instead of size
+  geom_point(size = 1.5) + # Point size
+  theme_minimal() + # Minimal theme
+  labs(title = "Crime Count by Type Over Years",
+         x = "Year",
+         y = "Number of Cases",
+         color = "Crime Type") +
+  theme(
+    plot.title = element_text(hjust = 0.5), # Center the plot title
+    legend.title = element_blank(), # Remove the legend title
+    axis.text.x = element_text(angle = 45, hjust = 1), # Angle the x-axis text for readability
+    text = element_text(size = 16), # Increase the global font size
+  )
 ```
@@
 
-## Data Visualization
-For visual representation, `ggplot2`, a versatile package in R for data visualization, was employed.
-It was used to create comprehensive line plots depicting the trends of various crime categories over the years.
-These plots are instrumental in visualizing the temporal patterns and changes in crime statistics.
+```{r}
+#| label: fig-trends-in-two-types-crime
+#| fig-cap: Comparative Trend Analysis of Violent and Non-Violent Crime Cases
+#| echo: false
 
-## Summary Statistics and Relationship Analysis
-The dataset was summarized using basic statistical measures (mean, median, standard deviation) to provide an overview of each crime category.
-Further, the relationships between different crime types were explored to understand correlated trends, if any.
+# Pivot the data to long format for plotting
+long_data <- two_types_crime %>%
+  select(Category, `2014`:`2023`) %>%
+  pivot_longer(cols = -Category, names_to = "Year", values_to = "Cases")
 
-# Results {#sec-result}
+# Create the plot
+ggplot(long_data, aes(x = Year, y = Cases, group = Category, color = Category)) +
+  geom_line(linewidth = 1) + # Use linewidth instead of size
+  geom_point(size = 2) + # Point size
+  theme_minimal() + # Minimal theme
+  labs(title = "Evolution of Violent and Non-Violent Crimes Over the Years",
+       x = "Year",
+       y = "Number of Cases") +
+  scale_color_manual(values = c("Violent" = "red", "Non-Violent" = "blue")) +
+  theme(
+    plot.title = element_text(hjust = 0.5), # Center the plot title
+    legend.title = element_blank(), # Remove the legend title
+    axis.text.x = element_text(angle = 45, hjust = 1) # Angle the x-axis text for readability
+  )
+```
+\newpage
 
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
+@fig-trends-in-two-types-crime compares the number of cases reported for violent and non-violent crimes from 2014 to 2023.
+The x-axis represents the years, and the y-axis indicates the number of cases.
 
-```
+The red line represents violent crimes, and the blue line represents non-violent crimes. The plot shows distinct trends for both categories over the specified time frame.
 
-```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
-#| warning: false
-```
+* **Violent Crimes (Red Line):** There is an overall increasing trend in the number of violent crime cases over the years.
+There are some fluctuations, with a noticeable dip occurring in one of the years, but the general direction is upwards, especially towards the end of the period, indicating a rise in violent crimes or an increase in reporting.
 
-Based on the analysis of the provided dataset and the visual representation of crime trends in Toronto over the years, several key results can be summarized:
+* **Non-Violent Crimes (Blue Line):** The trend in non-violent crime cases shows some variability but does not have as clear an increasing pattern as violent crimes.
+There are years where the rate has increased or decreased, but the last year shows a notable increase, suggesting a possible upward trend at the end of the period.
 
-* **Variation in Crime Trends:** The dataset reveals distinct trends for different types of crimes in Toronto.
-Some crime categories exhibit an upward trend, indicating an increase in incidents over the years, while others show a downward trend, suggesting a decrease in occurrences.
-This variation highlights the complexity of crime dynamics in an urban setting.
+Overall, the graph indicates that while both violent and non-violent crimes have seen rises and falls, the increase in violent crimes towards the later years is more pronounced.
+This could reflect changes in societal conditions, law enforcement practices, or reporting mechanisms.
 
-* **Potential Factors Influencing Trends:** The observed trends in crime rates could be influenced by a multitude of factors.
-These include socio-economic changes, urban development, law enforcement strategies, policy reforms, and technological advancements in crime reporting.
-For example, an increase in certain types of crimes might be linked to economic downturns, demographic shifts, or changes in urban density, while a decrease could reflect effective policing strategies or community intervention programs.
+\newpage
 
@@
 
-## Implications for Policymakers and Law Enforcement: {#sec-first-point}
+The analysis of Toronto's crime trends from 2014 to 2023 reveals a complex and evolving landscape of public safety.
+The pronounced increase in assault cases in recent years is alarming, suggesting that the city may be facing emerging challenges in maintaining public order.
+This rise in violent crime needs a closer investigation of possible causative factors, such as the effects of increased urbanization, economic disparity, and shifts in population density.
 
-The findings of this analysis are particularly relevant for policymakers and law enforcement agencies.
-Understanding these trends is crucial for developing targeted strategies to address and prevent crime effectively, thereby enhancing public safety and community well-being in Toronto.
+## Variation in Crime Trends: {#sec-first-point}
 
-## Second discussion point
+Simultaneously, the decline in theft-related crimes offers a more hopeful narrative.
+This trend may reflect the successful implementation of preventative measures by law enforcement agencies, such as community policing efforts and the integration of advanced surveillance technology.
+The use of crime prevention techniques, potentially aided by increased public awareness and education, appears to be bearing fruit.
+
+The contrast between the trajectories of violent and non-violent crime rates underscores the need for a differentiated approach to policy-making and law enforcement.
+While we appreciate our gains in decreasing property crimes, the data motivates us to pursue more effective interventions for preventing and reacting to violent crimes.
+This might include extending social services, improving community engagement, and investing in data-driven police strategies.
 
@@
 
-While the visual analysis provides a high-level overview of crime trends in Toronto, it's essential to conduct more detailed investigations to understand the underlying causes of these trends.
+While the visualization provides a high-level overview of crime trends in Toronto, it's essential to conduct more detailed analysis to understand the underlying causes of these trends.
 This includes examining subcategories of crimes, focusing on specific geographic areas within the city, and considering other relevant data sources.
+
+Furthermore, this study emphasises the need of ongoing monitoring and analysis of crime data in comprehending the ever-changing dynamics of urban crime.
+In light of these findings, future study should focus on understanding the influence of specific policies and social changes on crime rates.
+Only by doing such a comprehensive analysis can we hope to develop strategies that will ensure the safety and well-being of all Toronto residents.
 

### 2024-01-25 scripts/00-simulate_data.R

@@ -1,40 +1,38 @@
 #### Preamble ####
-# Purpose: Get data on 2021 shelter usage and make table
+# Purpose: Simulates data
 # Author: Yuanyi (Leo) Liu
 # Email: leoy.liu@mail.utoronto.ca
 # Date: 21 January 2024
+# License: MIT
+
 
 #### Workspace setup ####
-install.packages("knitr")
-install.packages("janitor")
-install.packages("lubridate")
-install.packages("opendatatoronto")
-install.packages("tidyverse")
-
-library(knitr)
+# install.packages("janitor")
+# install.packages("tidyverse")
 library(janitor)
-library(lubridate)
-library(opendatatoronto)
 library(tidyverse)
+
 
 #### Simulate ####
 
+# In this simulation, we only want the number of cases reported of each crime type from 2014 to 2023.
+# First, A tibble is created with initial Crime_Type column.
+# Second, A loop runs through each year from 2014 to 2023.
+# In each iteration of the loop, a new column is added to simulated_data for the respective year. 
+# The sample function generates random counts for the crimes for that year.
+# Third, replace = TRUE in the sample function allows for the possibility of the same number being chosen more than once.
+
 set.seed(853)
 
-simulated_occupancy_data <-
-  tibble(
-    date = rep(x = as.Date("2021-01-01") + c(0:364), times = 3),
-    # Based on Eddelbuettel: https://stackoverflow.com/a/21502386
-    shelter = c(
-      rep(x = "Shelter 1", times = 365),
-      rep(x = "Shelter 2", times = 365),
-      rep(x = "Shelter 3", times = 365)
-    ),
-    number_occupied =
-      rpois(
-        n = 365 * 3,
-        lambda = 30
-      ) # Draw 1,095 times from the Poisson distribution
-  )
+crime_types <- c("Assault", "Auto Theft", "Bike Theft", "Break and Enter", "Homicide", "Robbery", "Shooting Incidents", "Theft from Motor Vehicles", "Major Theft")
 
-head(simulated_occupancy_data)
+# Initialize a tibble
+simulated_data <- tibble(Crime_Type = crime_types)
+
+# Add columns for each year from 2014 to 2023 with simulated data
+for(year in 2014:2023) {
+  simulated_data[[as.character(year)]] <- sample(500:10000, length(crime_types), replace = TRUE)
+}
+
+# View the simulated data
+print(simulated_data)

### 2024-01-25 scripts/01-download_data.R

@@ -5,10 +5,12 @@
 # Contact: leoy.liu@mail.utoronto.ca
 # License: MIT
 
+
 #### Workspace setup ####
 library(opendatatoronto)
 library(tidyverse)
 library(dplyr)
+
 
 #### Download data ####
 # Each package is associated with a unique id  found in the "For 

### 2024-01-25 scripts/02-data_cleaning.R

@@ -5,10 +5,12 @@
 # Contact: leoy.liu@mail.utoronto.ca
 # License: MIT
 
+
 #### Workspace setup ####
 library(tidyverse)
 
-#### Clean data ####
+
+#### Cleaned data ####
 raw_data <- read_csv("inputs/data/toronto_crime_rates.csv")
 
 # Removing unwanted columns
@@ -17,12 +19,80 @@ cleaned_data <- raw_data %>% select(-c(X_id, HOOD_ID, POPULATION_2023, geometry)
 # Removing rows with NA values
 cleaned_data <- na.omit(cleaned_data)
 
-#### Save data ####
+# Save data #
 write_csv(
   x = cleaned_data,
   file = "cleaned_toronto_crime_rates.csv"
 )
 
-#### Descriptive Summaries ####
+# Descriptive Summaries #
 str(cleaned_data)
 summary(cleaned_data)
+
+
+
+#### Summary Table Dataset ####
+
+# Read the data
+crime_data <- read.csv(here("inputs/data/cleaned_toronto_crime_rates.csv"))
+
+# Remove unnecessary columns
+crime_data <- crime_data %>%
+  select(-grep("RATE", names(crime_data)))
+
+# Gather the data to a long format
+crime_data_long <- crime_data %>%
+  gather(key = "Crime_Year", value = "Count", -AREA_NAME)
+
+# Extract year and crime type from the 'Crime_Year' column
+crime_data_long <- crime_data_long %>%
+  separate(Crime_Year, into = c("Crime_Type", "Year"), sep = "_") %>%
+  mutate(Year = as.numeric(Year))
+
+# Aggregate the data by year and crime type
+crime_data_summary <- crime_data_long %>%
+  group_by(Crime_Type, Year) %>%
+  summarise(Total = sum(Count, na.rm = TRUE), .groups = "drop")
+
+# Creating the summary table
+summary_table <- crime_data_summary %>%
+  group_by(Crime_Type, Year) %>%
+  summarise(Total_Cases = sum(Total, na.rm = TRUE), .groups = "drop") %>%
+  spread(key = Year, value = Total_Cases)
+
+# Define a mapping for crime types
+readable_crime_types <- c('ASSAULT' = 'Assault',
+                          'AUTOTHEFT' = 'Auto Theft',
+                          'BIKETHEFT' = 'Bike Theft',
+                          'BREAKENTER' = 'Break and Enter',
+                          'HOMICIDE' = 'Homicide',
[…]

### 2024-01-25 outputs/paper/paper.qmd
finalize the paper

@@
 ## Data Manipulation and Cleaning
-The data was processed and cleaned using R[@citeR], a powerful statistical programming language. Key operations included:
-
-* **Data Cleaning:** The initial phase of cleaning the Toronto crime rates dataset involved importing the raw data using the `read_csv` function from the `readr`[@citereadr] package.
-Following the import, the dataset was streamlined using the `select` function from `dplyr`[@citedplyr], a tidyverse[@citetidyverse] package, to remove irrelevant columns such as `X_id`, `HOOD_ID`, `POPULATION_2023`, and `geometry`.
-This step was crucial to focus the analysis on relevant variables.
-Additionally, the `na.omit` function was employed to discard any rows with missing values (NA), ensuring the dataset's completeness and reliability.
-The final step in the cleaning process was the exportation of the cleaned data into a new CSV file using the `write_csv `function from `readr`[@citereadr], thereby preserving the cleaned and refined dataset for subsequent analysis.
-
-* **Data Transformation:** The dataset was transformed from a wide format to a long format using the `gather` function from the `tidyr`[@citetidyr] package.
-This transformation is crucial for simplifying the data structure and making it more amenable to analysis.
-Also, the separate function, again from `tidyr`[@citetidyr], was instrumental in dividing the 'Crime_Year' column into two distinct parts: 'Crime_Type' and 'Year'.
-This enhances the granularity of the dataset, allowing for more detailed analysis.
-
-* **Aggregation:** The `group_by` and `summarise` functions from `dplyr`[@citedplyr] were employed to aggregate the data by year and crime type.
-
-* **Data Visualization:** For visual representation, `ggplot2`[@citeggplot], a versatile package in R for data visualization, was employed.
-It was used to create comprehensive line plots depicting the trends of various crime categories over the years.
-
-* **Paper Generation:** The packages `knitr`[@citeknitr] is used in generating R Markdown reports.
+The data was processed and cleaned using R[@citeR], a powerful statistical programming language. For key operations please refer to the [@sec-appendix].
 
 ## Variables Explanation
+To better understanding the data, summary_table dataset was developed to offer a more structured and aggregated view of the crime data.
+By transforming and summarizing the data into a format that displays crime types against years, this dataset simplifies the task of identifying and analyzing trends over time.
 The dataset includes various crime categories, such as assault, theft, and other types, with each variable representing the annual count of reported cases for each crime type.
 This longitudinal data provides a detailed view of the crime trends in Toronto over a decade.
-
 ```{r}
@@
 We proceeded to organize the dataset into two categories: 'Violent' crimes, which included offenses such as 'Assault' and 'Homicide', and 'Non-Violent' crimes, encompassing acts like 'Theft'.
-Following this classification, we combined the data into these new groupings, totaling the number of incidents in each category for each year of the research.
+It is particularly valuable for understanding the overall landscape of crime in Toronto, focusing on the comparison between violent and non-violent incidents.
 ```{r}
@@
 
-## Variation in Crime Trends: {#sec-first-point}
+## Variation in Crime Trends {#sec-first-point}
 
@@
 # References
+
+::: {#refs}
+:::
+
+{{< pagebreak >}}
+
+# Appendix {#sec-appendix}
+## Data Manipulation and Cleaning
+* **Data Cleaning:** The initial phase of cleaning the Toronto crime rates dataset involved importing the raw data using the `read_csv` function from the `readr`[@citereadr] package.
+Following the import, the dataset was streamlined using the `select` function from `dplyr`[@citedplyr], a tidyverse[@citetidyverse] package, to remove irrelevant columns such as `X_id`, `HOOD_ID`, `POPULATION_2023`, and `geometry`.
+This step was crucial to focus the analysis on relevant variables.
+Additionally, the `na.omit` function was employed to discard any rows with missing values (NA), ensuring the dataset's completeness and reliability.
+The final step in the cleaning process was the exportation of the cleaned data into a new CSV file using the `write_csv `function from `readr`[@citereadr], thereby preserving the cleaned and refined dataset for subsequent analysis.
+
+* **Data Transformation:** The dataset was transformed from a wide format to a long format using the `gather` function from the `tidyr`[@citetidyr] package.
+This transformation is crucial for simplifying the data structure and making it more amenable to analysis.
+Also, the separate function, again from `tidyr`[@citetidyr], was instrumental in dividing the 'Crime_Year' column into two distinct parts: 'Crime_Type' and 'Year'.
+This enhances the granularity of the dataset, allowing for more detailed analysis.
+
+* **Aggregation:** The `group_by` and `summarise` functions from `dplyr`[@citedplyr] were employed to aggregate the data by year and crime type.
+
+* **Data Visualization:** For visual representation, `ggplot2`[@citeggplot], a versatile package in R for data visualization, was employed.
+It was used to create comprehensive line plots depicting the trends of various crime categories over the years.
+
+* **Paper Generation:** The packages `knitr`[@citeknitr] is used in generating R Markdown reports.

### 2024-01-25 outputs/paper/paper.qmd
check for grammar

@@
 The findings of this paper are noteworthy for several reasons.
-First, they provide a nuanced understanding of the crime dynamics a large metropolitan area, offering insights that are vital for policymakers, law enforcement, and community leaders.
+First, they provide a nuanced understanding of the crime dynamics in a large metropolitan area, offering insights that are vital for policymakers, law enforcement, and community leaders.
 The patterns observed suggest shifts in criminal behavior and possibly reflect broader social and economic changes in the city.
@@
 @sec-result discusses the results, elaborating on the observed trends and patterns in crime data.
-@sec-discussion provides an in-depth discussion of these findings, exploring potential factors influencing these trends, drawing connections to broader urban issues, and suggestions for future research in this area.
+@sec-discussion provides an in-depth discussion of these findings, exploring potential factors influencing these trends, drawing connections to broader urban issues, and providing suggestions for future research in this area.
 
@@
 ## Data Manipulation and Cleaning
-The data was processed and cleaned using R[@citeR], a powerful statistical programming language. For key operations please refer to the [@sec-appendix].
+The data was processed and cleaned using R[@citeR], a powerful statistical programming language. For key operations, please refer to the [@sec-appendix].
 
 ## Variables Explanation
-To better understanding the data, summary_table dataset was developed to offer a more structured and aggregated view of the crime data.
+To better understand the data, the summary_table dataset was developed to offer a more structured and aggregated view of the crime data.
 By transforming and summarizing the data into a format that displays crime types against years, this dataset simplifies the task of identifying and analyzing trends over time.
@@
 # Read the data
-summary_table <- read.csv(here("inputs/data/summary_table.csv"), check.names = FALSE)
+summary_table <- read.csv(here("outputs/data/summary_table.csv"), check.names = FALSE)
 
@@
 # Read the data
-two_types_crime <- read.csv(here("inputs/data/two_types_crime.csv"), check.names = FALSE)
+two_types_crime <- read.csv(here("outputs/data/two_types_crime.csv"), check.names = FALSE)
 
@@
 @fig-trends-in-crime reveals distinct trends for various crime types in Toronto over the years.
-We notice that assault crimes have been on the rise in recent years, this could signal a growing concern in public safety and might prompt law enforcement to investigate the underlying causes, such as increased urbanization, changes in population density, or other socio-economic factors.
-Conversely, there is a noticeable decline in theft-related crimes, it could reflect the success of recent crime prevention initiatives, improved security measures, or increased public awareness.
+We notice that assault crimes have been on the rise in recent years.
+This could signal a growing concern in public safety and might prompt law enforcement to investigate the underlying causes, such as increased urbanization, changes in population density, or other socio-economic factors.
+Conversely, there is a noticeable decline in theft-related crimes, which could reflect the success of recent crime prevention initiatives, improved security measures, or increased public awareness.
 
@@
 
-* **Violent Crimes (Red Line):** There is an overall increasing trend in the number of violent crime cases over the years.
+* **Violent Crimes (Red Line):** There has been an overall increasing trend in the number of violent crime cases over the years.
 There are some fluctuations, with a noticeable dip occurring in one of the years, but the general direction is upwards, especially towards the end of the period, indicating a rise in violent crimes or an increase in reporting.
@@
 
-## Weaknesses and next steps
+## Weaknesses and Next Steps
 
-While the visualization provides a high-level overview of crime trends in Toronto, it's essential to conduct more detailed analysis to understand the underlying causes of these trends.
+While the visualization provides a high-level overview of crime trends in Toronto, it's essential to conduct a more detailed analysis to understand the underlying causes of these trends.
 This includes examining subcategories of crimes, focusing on specific geographic areas within the city, and considering other relevant data sources.
 
-Furthermore, this study emphasises the need of ongoing monitoring and analysis of crime data in comprehending the ever-changing dynamics of urban crime.
-In light of these findings, future study should focus on understanding the influence of specific policies and social changes on crime rates.
+Furthermore, this study emphasises the need for ongoing monitoring and analysis of crime data to comprehend the ever-changing dynamics of urban crime.
+In light of these findings, future studies should focus on understanding the influence of specific policies and social changes on crime rates.
 Only by doing such a comprehensive analysis can we hope to develop strategies that will ensure the safety and well-being of all Toronto residents.

### 2024-01-25 outputs/paper/paper.qmd
change to correct github url

@@
   - Yuanyi (Leo) Liu
-thanks: "Code and data are available at: https://github.com/leoyliu/STA302-Paper-1."
+thanks: "Code and data are available at: https://github.com/leoyliu/Decoding-Toronto-Crime-Trends-from-2014-to-2023."
 date: today
